---
title: "Bayesian Modelling"
author: "Kenéz"
date: "2025-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# seed was 101git branch -D main
set.seed(10101)

library(dplyr)
library(brms)
library(ggplot2)
library(tidyr)
library(glue) 
library(caret)
library(tidyverse)
library(randomForest)
library(caTools)
library(bayesplot)
library(glmnet)
library(here)
library(priorsense)
```

In this notebook I will make some Bayesian models and compare them, to hopefully establish a set of good modelling choices.

1 -- I test whether Gelman priors on non-game predictors are fitting or are too informative.

```{r}
# Specifying formula and priors
formula <- bf(Dyslexia ~ Female + Monolingual + Failedclass + Age)

sd_Dyslexia <- sd(training_scaled$Dyslexia)
sd_Female <- sd(training_scaled$Female)
sd_Monolingual <- sd(training_scaled$Monolingual)
sd_Failedclass <- sd(training_scaled$Failedclass)
sd_Age <- sd(training_scaled$Age)


priors <- c(
  prior(normal(-2, 1), class = Intercept),
  prior(normal(0, 2.5*0.3109777/0.9998264), class = b, coef = Female),
  prior(normal(0, 2.5*0.3109777/0.8841844), class = b, coef = Monolingual),
  prior(normal(0, 2.5*0.3109777/0.7845354), class = b, coef = Failedclass),
  prior(normal(0, 2.5*0.3109777/1), class = b, coef = Age)
#  prior(exponential(1/sd(training_scaled$Dyslexia)), class = sigma)
)

```

```{r message = False }
# Specifying, running the model

start_time <- Sys.time()

gelman_prior_check_model <- brm(formula = formula,
           prior = priors,
           family = bernoulli(link = 'logit'),
           data = training_scaled
           )
end_time <- Sys.time()
runtime <- end_time - start_time

```
```{r}
# Seeing if there are any priors that are too strong.

powerscale_sensitivity(gelman_prior_check_model)

```
Even though there the priorsense package warns of prior-data conflicts, I feel the priors are correct. We strongly suspect around 10% chance of dyslexia in each participant. The variance is high enough anyways.


```{r}

# Specifying formula
formula <- bf(Dyslexia ~ a + b, nl = TRUE) +
  lf(a ~ Clicks1 + Hits1 + Misses1 + Score1 + Accuracy1 + Missrate1 + Clicks2 + Hits2 + Misses2 + Score2 + Accuracy2 + Missrate2 + Clicks3 + Hits3 + Misses3 + Score3 + Accuracy3 + Missrate3 + Clicks4 + Hits4 + Misses4 + Score4 + Accuracy4 + Missrate4 + Clicks5 + Hits5 + Misses5 + Score5 + Accuracy5 + Missrate5 + Clicks6 + Hits6 + Misses6 + Score6 + Accuracy6 + Missrate6 + Clicks7 + Hits7 + Misses7 + Score7 + Accuracy7 + Missrate7 + Clicks8 + Hits8 + Misses8 + Score8 + Accuracy8 + Missrate8 + Clicks9 + Hits9 + Misses9 + Score9 + Accuracy9 + Missrate9 + Clicks10 + Hits10 + Misses10 + Score10 + Accuracy10 + Missrate10 + Clicks11 + Hits11 + Misses11 + Score11 + Accuracy11 + Missrate11 + Clicks12 + Hits12 + Misses12 + Score12 + Accuracy12 + Missrate12 + Clicks13 + Hits13 + Misses13 + Score13 + Accuracy13 + Missrate13 + Clicks14 + Hits14 + Misses14 + Score14 + Accuracy14 + Missrate14 + Clicks15 + Hits15 + Misses15 + Score15 + Accuracy15 + Missrate15 + Clicks16 + Hits16 + Misses16 + Score16 + Accuracy16 + Missrate16 + Clicks17 + Hits17 + Misses17 + Score17 + Accuracy17 + Missrate17 + Clicks18 + Hits18 + Misses18 + Score18 + Accuracy18 + Missrate18 + Clicks19 + Hits19 + Misses19 + Score19 + Accuracy19 + Missrate19 + Clicks20 + Hits20 + Misses20 + Score20 + Accuracy20 + Missrate20 + Clicks21 + Hits21 + Misses21 + Score21 + Accuracy21 + Missrate21 + Clicks22 + Hits22 + Misses22 + Score22 + Accuracy22 + Missrate22 + Clicks23 + Hits23 + Misses23 + Score23 + Accuracy23 + Missrate23 + Clicks24 + Hits24 + Misses24 + Score24 + Accuracy24 + Missrate24 + Clicks25 + Hits25 + Misses25 + Score25 + Accuracy25 + Missrate25 + Clicks26 + Hits26 + Misses26 + Score26 + Accuracy26 + Missrate26 + Clicks27 + Hits27 + Misses27 + Score27 + Accuracy27 + Missrate27 + Clicks28 + Hits28 + Misses28 + Score28 + Accuracy28 + Missrate28 + Clicks29 + Hits29 + Misses29 + Score29 + Accuracy29 + Missrate29 + Clicks30 + Hits30 + Misses30 + Score30 + Accuracy30 + Missrate30 + Clicks31 + Hits31 + Misses31 + Score31 + Accuracy31 + Missrate31 + Clicks32 + Hits32 + Misses32 + Score32 + Accuracy32, center = TRUE) +
  lf(b ~ 0 + Nativelang + Otherlang + Age + Female, cmc = FALSE)

# Specifying priors
priors <- prior(horseshoe(1), class = "b", nlpar = "a")

# Specifying the model
start_time <- Sys.time()

bayesian_model <- brm(formula = formula,
           prior = priors,
           family = bernoulli(link = 'logit'),
           data = training
           )

end_time <- Sys.time()
runtime <- end_time - start_time

writeLines(runtime, "runtime.txt")
saveRDS(bayesian_model, file = "bayesian_model.rds")
```



