---
title: "Bayesian Modelling"
author: "Kenéz"
date: "2025-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(1337)

library(dplyr)
library(brms)
library(ggplot2)
library(tidyr)
library(glue) 
library(caret)
library(tidyverse)
library(randomForest)
library(caTools)
library(bayesplot)
library(glmnet)
library(here)
library(priorsense)

log_file_with_path <- here::here("results", "runtime_log.txt")

source(here::here("scripts", "functions.R"))
```

In this notebook I will make some Bayesian models and compare them, to hopefully establish a set of good modelling choices. I will use the decision threshold 0.24.

1 -- I test whether Gelman priors on non-game predictors are fitting or are too informative.

```{r}
# Specifying formula and priors
formula <- bf(Dyslexia ~ Female + Monolingual + Failedclass + Age)

sd_Dyslexia <- sd(training_scaled$Dyslexia)
sd_Female <- sd(training_scaled$Female)
sd_Monolingual <- sd(training_scaled$Monolingual)
sd_Failedclass <- sd(training_scaled$Failedclass)
sd_Age <- sd(training_scaled$Age)


priors <- c(
  prior(normal(-2, 1), class = Intercept),
  prior(normal(0, 2.5*0.3109777/0.9998264), class = b, coef = Female),
  prior(normal(0, 2.5*0.3109777/0.8841844), class = b, coef = Monolingual),
  prior(normal(0, 2.5*0.3109777/0.7845354), class = b, coef = Failedclass),
  prior(normal(0, 2.5*0.3109777/1), class = b, coef = Age)
#  prior(exponential(1/sd(training_scaled$Dyslexia)), class = sigma)
)

```

```{r message = False }
# Specifying, running the model
#TODO: Make runtime be saved in a file? a nicely accessible excel file would be nice. or something like that
start_time <- Sys.time()

model_without_gamevariables <- brm(formula = formula,
           prior = priors,
           family = bernoulli(link = 'logit'),
           data = training_scaled
           )

end_time <- Sys.time()
log_runtime(start_time, end_time, "model_without_gamevariables", log_file_with_path)

saveRDS(bayesian_model, file = here("models", "bayesian_model.rds"))
```

```{r}
# Seeing if there are any priors that are too strong.

powerscale_sensitivity(gelman_prior_check_model)
```

No prior-data conflicts, were all good. :)

2 -- The game variables part.

Now that we have our non-game variables in good order, outfitted with nice priors, we can start trying to include the game variables as well.

```{r}

# Specifying formula
formula <- bf(Dyslexia ~ a + b, nl = TRUE) +
  lf(a ~ Clicks1 + Hits1 + Misses1 + Score1 + Accuracy1 + Missrate1 + Clicks2 + Hits2 + Misses2 + Score2 + Accuracy2 + Missrate2 + Clicks3 + Hits3 + Misses3 + Score3 + Accuracy3 + Missrate3 + Clicks4 + Hits4 + Misses4 + Score4 + Accuracy4 + Missrate4 + Clicks5 + Hits5 + Misses5 + Score5 + Accuracy5 + Missrate5 + Clicks6 + Hits6 + Misses6 + Score6 + Accuracy6 + Missrate6 + Clicks7 + Hits7 + Misses7 + Score7 + Accuracy7 + Missrate7 + Clicks8 + Hits8 + Misses8 + Score8 + Accuracy8 + Missrate8 + Clicks9 + Hits9 + Misses9 + Score9 + Accuracy9 + Missrate9 + Clicks10 + Hits10 + Misses10 + Score10 + Accuracy10 + Missrate10 + Clicks11 + Hits11 + Misses11 + Score11 + Accuracy11 + Missrate11 + Clicks12 + Hits12 + Misses12 + Score12 + Accuracy12 + Missrate12 + Clicks13 + Hits13 + Misses13 + Score13 + Accuracy13 + Missrate13 + Clicks14 + Hits14 + Misses14 + Score14 + Accuracy14 + Missrate14 + Clicks15 + Hits15 + Misses15 + Score15 + Accuracy15 + Missrate15 + Clicks16 + Hits16 + Misses16 + Score16 + Accuracy16 + Missrate16 + Clicks17 + Hits17 + Misses17 + Score17 + Accuracy17 + Missrate17 + Clicks18 + Hits18 + Misses18 + Score18 + Accuracy18 + Missrate18 + Clicks19 + Hits19 + Misses19 + Score19 + Accuracy19 + Missrate19 + Clicks20 + Hits20 + Misses20 + Score20 + Accuracy20 + Missrate20 + Clicks21 + Hits21 + Misses21 + Score21 + Accuracy21 + Missrate21 + Clicks22 + Hits22 + Misses22 + Score22 + Accuracy22 + Missrate22 + Clicks23 + Hits23 + Misses23 + Score23 + Accuracy23 + Missrate23 + Clicks24 + Hits24 + Misses24 + Score24 + Accuracy24 + Missrate24 + Clicks25 + Hits25 + Misses25 + Score25 + Accuracy25 + Missrate25 + Clicks26 + Hits26 + Misses26 + Score26 + Accuracy26 + Missrate26 + Clicks27 + Hits27 + Misses27 + Score27 + Accuracy27 + Missrate27 + Clicks28 + Hits28 + Misses28 + Score28 + Accuracy28 + Missrate28 + Clicks29 + Hits29 + Misses29 + Score29 + Accuracy29 + Missrate29 + Clicks30 + Hits30 + Misses30 + Score30 + Accuracy30 + Missrate30 + Clicks31 + Hits31 + Misses31 + Score31 + Accuracy31 + Missrate31 + Clicks32 + Hits32 + Misses32 + Score32 + Accuracy32, center = TRUE) +
  lf(b ~ 0 + Monolingual + Failedclass + Age + Female, cmc = FALSE)

# Specifying priors 
#TODO: intercept prior normal(-2,1)
priors <- c(
  prior(horseshoe(1), class = "b", nlpar = "a"),
  prior(normal(0, 2.5*0.3109777/0.9998264), coef = "Female", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/0.8841844), coef = "Monolingual", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/0.7845354), coef = "Failedclass", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/1), coef = "Age", nlpar = "b")
)

# Specifying the model
start_time <- Sys.time()

bayesian_model <- brm(formula = formula,
           prior = priors,
           family = bernoulli(link = 'logit'),
           data = training
           )

end_time <- Sys.time()

log_runtime(start_time, end_time, "m_allvariables_horseshoe_gelman", log_file_with_path)

saveRDS(bayesian_model, file = here("models", "bayesian_model.rds"))
```

The chunk below makes a confusion matrix for the model in question.

```{r echo = False}
model <- bayesian_model

# Predict probabilities using `fitted`
model_predicted_probabilities <- fitted(model, newdata = validation_scaled)
predicted_probs <- model_predicted_probabilities[, 1]  # Extract the mean probabilities

# Convert probabilities to binary predictions using a threshold (e.g., 0.24)
predicted_classes <- ifelse(predicted_probs > 0.24, 1, 0)

# Ensure `predicted_classes` and `validation$Dyslexia_binary` are factors with the same levels
predicted_classes <- factor(predicted_classes, levels = c(0, 1))
actual_classes <- factor(validation_scaled$Dyslexia, levels = c(0, 1))

# Check lengths to make sure they match
if (length(predicted_classes) != length(actual_classes)) {
  stop("Length mismatch: Predicted length =", length(predicted_classes), 
       "Actual length =", length(actual_classes))
}

# Create a confusion matrix
library(caret)
cm <- confusionMatrix(predicted_classes, actual_classes)

# Print the confusion matrix
print(cm)
```
3 -- Comparing the model with and without the game variables

I'd like to know how useful the game variables are. To this end, I will compare two models with loo; one with the game variables and one without. The difference between their predictive performance will give us an idea as to how useful the game is.

```{r}
# Getting loo for each model
loo_model1 <- loo(model_without_gamevariables)
loo_model2 <- loo(bayesian_model)

# Compare models using LOO
comparison <- loo_compare(loo_model1, loo_model2)

# Print the comparison
print(comparison)
```
The bayesian model seems to do much better, so the game variables help a lot. :)

4 -- Trying to get the model to perform better

We've established that the model with game variables performs better. Therefore, our modelling paradigm can capture the variance there to some extent. However, apparently, a random forest would do better. Our objective therefore is to find a way to manipulate either the model specifications or the data in order to make the model perform better.

I'll first start by doing some data manipulation and seeing if that makes a difference or not.

4.1 -- Dropping misses

Misses are perfectly explained by Hits and Clicks in theory. In reality somehow they are not, in our data.

```{r}

# Specifying formula
formula <- bf(Dyslexia ~ a + b, nl = TRUE) +
  lf(a ~ Clicks1 + Hits1 + Score1 + Accuracy1 + Missrate1 + Clicks2 + Hits2 + Score2 + Accuracy2 + Missrate2 + Clicks3 + Hits3 + Score3 + Accuracy3 + Missrate3 + Clicks4 + Hits4 + Score4 + Accuracy4 + Missrate4 + Clicks5 + Hits5 + Score5 + Accuracy5 + Missrate5 + Clicks6 + Hits6 + Score6 + Accuracy6 + Missrate6 + Clicks7 + Hits7 + Score7 + Accuracy7 + Missrate7 + Clicks8 + Hits8 + Score8 + Accuracy8 + Missrate8 + Clicks9 + Hits9 + Score9 + Accuracy9 + Missrate9 + Clicks10 + Hits10 + Score10 + Accuracy10 + Missrate10 + Clicks11 + Hits11 + Score11 + Accuracy11 + Missrate11 + Clicks12 + Hits12 + Score12 + Accuracy12 + Missrate12 + Clicks13 + Hits13 + Score13 + Accuracy13 + Missrate13 + Clicks14 + Hits14 + Score14 + Accuracy14 + Missrate14 + Clicks15 + Hits15 + Score15 + Accuracy15 + Missrate15 + Clicks16 + Hits16 + Score16 + Accuracy16 + Missrate16 + Clicks17 + Hits17 + Score17 + Accuracy17 + Missrate17 + Clicks18 + Hits18 + Score18 + Accuracy18 + Missrate18 + Clicks19 + Hits19 + Score19 + Accuracy19 + Missrate19 + Clicks20 + Hits20 + Score20 + Accuracy20 + Missrate20 + Clicks21 + Hits21 + Score21 + Accuracy21 + Missrate21 + Clicks22 + Hits22 + Score22 + Accuracy22 + Missrate22 + Clicks23 + Hits23 + Score23 + Accuracy23 + Missrate23 + Clicks24 + Hits24 + Score24 + Accuracy24 + Missrate24 + Clicks25 + Hits25 + Score25 + Accuracy25 + Missrate25 + Clicks26 + Hits26 + Score26 + Accuracy26 + Missrate26 + Clicks27 + Hits27 + Score27 + Accuracy27 + Missrate27 + Clicks28 + Hits28 + Score28 + Accuracy28 + Missrate28 + Clicks29 + Hits29 + Score29 + Accuracy29 + Missrate29 + Clicks30 + Hits30 + Score30 + Accuracy30 + Missrate30 + Clicks31 + Hits31 + Score31 + Accuracy31 + Missrate31 + Clicks32 + Hits32 + Score32 + Accuracy32 + Missrate32, center = TRUE) +
lf(b ~ 0 + Monolingual + Failedclass + Age + Female, cmc = FALSE)

# Specifying priors 
#TODO: intercept prior normal(-2,1)
priors <- c(
  prior(horseshoe(1), class = "b", nlpar = "a"),
  prior(normal(0, 2.5*0.3109777/0.9998264), coef = "Female", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/0.8841844), coef = "Monolingual", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/0.7845354), coef = "Failedclass", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/1), coef = "Age", nlpar = "b")
)

# Specifying the model
start_time <- Sys.time()

bayesian_model_without_misses <- brm(formula = formula,
           prior = priors,
           family = bernoulli(link = 'logit'),
           data = training
           )

end_time <- Sys.time()

log_runtime(start_time, end_time, "m_allvariables_horseshoe_gelman_missesremoved", log_file_with_path)

saveRDS(bayesian_model, file = here("models", "m_allvariables_horseshoe_gelman_missesremoved.rds"))
```

```{r, echo = FALSE}
# Getting loo for each model
loo_model1 <- loo(bayesian_model_without_misses)
loo_model2 <- loo(bayesian_model)

# Compare models using LOO
comparison <- loo_compare(loo_model1, loo_model2)

# Print the comparison
print(comparison)
```

Conclusion: Dropping misses is not a good idea. There is some new information there.

4.2 -- Dropping missrate

Missrate is perfectly explained by Accuracy.

```{r}

# Specifying formula
formula <- bf(Dyslexia ~ a + b, nl = TRUE) +
  lf(a ~ Clicks1 + Hits1 + Misses1 + Score1 + Accuracy1 + Clicks2 + Hits2 + Misses2 + Score2 + Accuracy2 + Clicks3 + Hits3 + Misses3 + Score3 + Accuracy3 + Clicks4 + Hits4 + Misses4 + Score4 + Accuracy4 + Clicks5 + Hits5 + Misses5 + Score5 + Accuracy5 + Clicks6 + Hits6 + Misses6 + Score6 + Accuracy6 + Clicks7 + Hits7 + Misses7 + Score7 + Accuracy7 + Clicks8 + Hits8 + Misses8 + Score8 + Accuracy8 + Clicks9 + Hits9 + Misses9 + Score9 + Accuracy9 + Clicks10 + Hits10 + Misses10 + Score10 + Accuracy10 + Clicks11 + Hits11 + Misses11 + Score11 + Accuracy11 + Clicks12 + Hits12 + Misses12 + Score12 + Accuracy12 + Clicks13 + Hits13 + Misses13 + Score13 + Accuracy13 + Clicks14 + Hits14 + Misses14 + Score14 + Accuracy14 + Clicks15 + Hits15 + Misses15 + Score15 + Accuracy15 + Clicks16 + Hits16 + Misses16 + Score16 + Accuracy16 + Clicks17 + Hits17 + Misses17 + Score17 + Accuracy17 + Clicks18 + Hits18 + Misses18 + Score18 + Accuracy18 + Clicks19 + Hits19 + Misses19 + Score19 + Accuracy19 + Clicks20 + Hits20 + Misses20 + Score20 + Accuracy20 + Clicks21 + Hits21 + Misses21 + Score21 + Accuracy21 + Clicks22 + Hits22 + Misses22 + Score22 + Accuracy22 + Clicks23 + Hits23 + Misses23 + Score23 + Accuracy23 + Clicks24 + Hits24 + Misses24 + Score24 + Accuracy24 + Clicks25 + Hits25 + Misses25 + Score25 + Accuracy25 + Clicks26 + Hits26 + Misses26 + Score26 + Accuracy26 + Clicks27 + Hits27 + Misses27 + Score27 + Accuracy27 + Clicks28 + Hits28 + Misses28 + Score28 + Accuracy28 + Clicks29 + Hits29 + Misses29 + Score29 + Accuracy29 + Clicks30 + Hits30 + Misses30 + Score30 + Accuracy30 + Clicks31 + Hits31 + Misses31 + Score31 + Accuracy31 + Clicks32 + Hits32 + Misses32 + Score32 + Accuracy32, center = TRUE) +
lf(b ~ 0 + Monolingual + Failedclass + Age + Female, cmc = FALSE)

# Specifying priors 
#TODO: intercept prior normal(-2,1)
priors <- c(
  prior(horseshoe(1), class = "b", nlpar = "a"),
  prior(normal(0, 2.5*0.3109777/0.9998264), coef = "Female", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/0.8841844), coef = "Monolingual", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/0.7845354), coef = "Failedclass", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/1), coef = "Age", nlpar = "b")
)

# Specifying the model
start_time <- Sys.time()

bayesian_model_without_missrates <- brm(formula = formula,
           prior = priors,
           family = bernoulli(link = 'logit'),
           data = training
           )

end_time <- Sys.time()

log_runtime(start_time, end_time, "m_allvariables_horseshoe_gelman_missratesremoved", log_file_with_path)

saveRDS(bayesian_model_without_missrates, file = here("models", "m_allvariables_horseshoe_gelman_missratesremoved.rds"))
```

```{r, echo = FALSE}
# Getting loo for each model
loo_model1 <- loo(bayesian_model_without_missrates)
loo_model2 <- loo(bayesian_model)

# Compare models using LOO
comparison <- loo_compare(loo_model1, loo_model2)

# Print the comparison
print(comparison)
```

Conclusion: Excluding missrates makes sense.

4.3 - Excluding Score

Score is probably just a linear combination of other recorded game variables.


```{r}

# Specifying formula
formula <- bf(Dyslexia ~ a + b, nl = TRUE) +
  lf(a ~ Clicks1 + Hits1 + Misses1 + Accuracy1 + Missrate1 +
    Clicks2 + Hits2 + Misses2 + Accuracy2 + Missrate2 +
    Clicks3 + Hits3 + Misses3 + Accuracy3 + Missrate3 +
    Clicks4 + Hits4 + Misses4 + Accuracy4 + Missrate4 +
    Clicks5 + Hits5 + Misses5 + Accuracy5 + Missrate5 +
    Clicks6 + Hits6 + Misses6 + Accuracy6 + Missrate6 +
    Clicks7 + Hits7 + Misses7 + Accuracy7 + Missrate7 +
    Clicks8 + Hits8 + Misses8 + Accuracy8 + Missrate8 +
    Clicks9 + Hits9 + Misses9 + Accuracy9 + Missrate9 +
    Clicks10 + Hits10 + Misses10 + Accuracy10 + Missrate10 +
    Clicks11 + Hits11 + Misses11 + Accuracy11 + Missrate11 +
    Clicks12 + Hits12 + Misses12 + Accuracy12 + Missrate12 +
    Clicks13 + Hits13 + Misses13 + Accuracy13 + Missrate13 +
    Clicks14 + Hits14 + Misses14 + Accuracy14 + Missrate14 +
    Clicks15 + Hits15 + Misses15 + Accuracy15 + Missrate15 +
    Clicks16 + Hits16 + Misses16 + Accuracy16 + Missrate16 +
    Clicks17 + Hits17 + Misses17 + Accuracy17 + Missrate17 +
    Clicks18 + Hits18 + Misses18 + Accuracy18 + Missrate18 +
    Clicks19 + Hits19 + Misses19 + Accuracy19 + Missrate19 +
    Clicks20 + Hits20 + Misses20 + Accuracy20 + Missrate20 +
    Clicks21 + Hits21 + Misses21 + Accuracy21 + Missrate21 +
    Clicks22 + Hits22 + Misses22 + Accuracy22 + Missrate22 +
    Clicks23 + Hits23 + Misses23 + Accuracy23 + Missrate23 +
    Clicks24 + Hits24 + Misses24 + Accuracy24 + Missrate24 +
    Clicks25 + Hits25 + Misses25 + Accuracy25 + Missrate25 +
    Clicks26 + Hits26 + Misses26 + Accuracy26 + Missrate26 +
    Clicks27 + Hits27 + Misses27 + Accuracy27 + Missrate27 +
    Clicks28 + Hits28 + Misses28 + Accuracy28 + Missrate28 +
    Clicks29 + Hits29 + Misses29 + Accuracy29 + Missrate29 +
    Clicks30 + Hits30 + Misses30 + Accuracy30 + Missrate30 +
    Clicks31 + Hits31 + Misses31 + Accuracy31 + Missrate31 +
    Clicks32 + Hits32 + Misses32 + Accuracy32 + Missrate32,
    center = TRUE) +
lf(b ~ 0 + Monolingual + Failedclass + Age + Female, cmc = FALSE)

# Specifying priors 
#TODO: intercept prior normal(-2,1)
priors <- c(
  prior(horseshoe(1), class = "b", nlpar = "a"),
  prior(normal(0, 2.5*0.3109777/0.9998264), coef = "Female", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/0.8841844), coef = "Monolingual", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/0.7845354), coef = "Failedclass", nlpar = "b"),
  prior(normal(0, 2.5*0.3109777/1), coef = "Age", nlpar = "b")
)

# Specifying the model
start_time <- Sys.time()

bayesian_model_without_scores <- brm(formula = formula,
           prior = priors,
           family = bernoulli(link = 'logit'),
           data = training
           )

end_time <- Sys.time()

log_runtime(start_time, end_time, "m_allvariables_horseshoe_gelman_scoresremoved", log_file_with_path)

saveRDS(bayesian_model_without_scores, file = here("models", "m_allvariables_horseshoe_gelman_scoresremoved.rds"))
```

```{r, echo = FALSE}
# Getting loo for each model
loo_model1 <- loo(bayesian_model_without_scores)
loo_model2 <- loo(bayesian_model)

# Compare models using LOO
comparison <- loo_compare(loo_model1, loo_model2)

# Print the comparison
print(comparison)
```

Conclusion: Dropping score is not a good idea? Its very close though.