---
title: "Data Cleaning"
author: "Kenéz"
date: "2025-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# seed was 101git branch -D main
set.seed(10101)

library(dplyr)
library(brms)
library(ggplot2)
library(tidyr)
library(glue) 
library(caret)
library(tidyverse)
library(randomForest)
library(caTools)
library(bayesplot)
library(glmnet)
library(here)
```
## Data

Two sources of data are available: The original data, collected through the use of computers, and a second source, collected on tablets with a modified game. (The game had less questions.) No attempt has been made yet to use both datasets; I abandoned my efforts after preliminary testing showed no real difference in performance when I included the secondary dataset. However, I think with a bit more work, perhaps a better imputation method, I could find a way to exploit the information contained in the secondary dataset.

```{r}
data <- read.csv(here("data", "Dyt-desktop.csv"), sep=";", dec=".", header=TRUE)

# Accuracy and Missrate are not accurate in many cases; I fix it here
for (i in 1:32) {
  accuracy_col <- paste0("Accuracy", i)
  missrate_col <- paste0("Missrate", i)
  hits_col <- paste0("Hits", i)
  clicks_col <- paste0("Clicks", i)
  misses_col <- paste0("Misses", i)
  data[[accuracy_col]] <- ifelse(data[[clicks_col]] == 0, 0, data[[hits_col]] / data[[clicks_col]])
  data[[missrate_col]] <- ifelse(data[[clicks_col]] == 0, 00, data[[misses_col]] / data[[clicks_col]])
}

# Changing Gender column to Female column
data$Female <- ifelse(data$Gender == "Male", "Yes", "No")
data$Gender <- NULL

# Factorizing the categorical variables
data$Female <- factor(data$Female, levels = c("No", "Yes"))
data$Nativelang <- factor(data$Nativelang, levels = c("No", "Yes"))
data$Otherlang <- factor(data$Otherlang, levels = c("No", "Yes"))
data$Dyslexia <- factor(data$Dyslexia, levels = c("No", "Yes"))

# Train-test split
sample = sample.split(data[,1], SplitRatio = 0.8)
traintemp = subset(data, sample == TRUE) # this is training and validation, before their split
testing  = subset(data, sample == FALSE)

# Train-validation split
sample = sample.split(traintemp[,1], SplitRatio = 0.75)
training = subset(traintemp, sample == TRUE)
validation  = subset(traintemp, sample == FALSE)
```

Here is how the data looks like:
```{r echo = FALSE}
head(data[, 1:12], 3)
```

```{r echo = FALSE}
# Histogram for 'Age' column
ggplot(data, aes(x = Age)) + 
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Age", x = "Age", y = "Count") +
  theme_minimal()

# Count the cases of Dyslexia = "Yes"
dyslexia_yes <- sum(data$Dyslexia == "Yes")
dyslexia_no <- sum(data$Dyslexia == "No")
cat("Number of children with Dyslexia:", dyslexia_yes, "\n")
cat("Number of children without Dyslexia:", dyslexia_no, "\n")
```
This is the code I used to standardise. In the thesis I did not use standardised data; I could not get better model results. This is strange since one would think the shrinkage priors would do better with standardised data. 

```{r}
exclude_columns <- c("Female", "Nativelang", "Otherlang", "Tablet", "Dyslexia")
columns_to_standardize <- setdiff(names(training), exclude_columns)

# Using the mean and sd of the training subset to not get data leakage
train_mean <- sapply(training[, columns_to_standardize], mean)
train_sd <- sapply(training[, columns_to_standardize], sd)

# Standardising the training, validation and testing data
training_scaled <- training
training_scaled[, columns_to_standardize] <- scale(training[, columns_to_standardize], center = train_mean, scale = train_sd)

validation_scaled <- validation
validation_scaled[, columns_to_standardize] <- scale(validation[, columns_to_standardize], center = train_mean, scale = train_sd)

testing_scaled <- testing
testing_scaled[, columns_to_standardize] <- scale(testing[, columns_to_standardize], center = train_mean, scale = train_sd)
```

```{r}
write.csv(data, here("data", "data.csv"), row.names = FALSE)
write.csv(traintemp, here("data", "traintemp.csv"), row.names = FALSE)
write.csv(training, here("data", "training.csv"), row.names = FALSE)
write.csv(testing, here("data", "testing.csv"), row.names = FALSE)
write.csv(validation, here("data", "validation.csv"), row.names = FALSE)
write.csv(training_scaled, here("data", "training_scaled.csv"), row.names = FALSE)
write.csv(testing_scaled, here("data", "testing_scaled.csv"), row.names = FALSE)
write.csv(validation_scaled, here("data", "validation_scaled.csv"), row.names = FALSE)
```