---
title: "Dyslexia_Paper_Code"
author: "Kenéz"
date: "2025-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# seed was 101
set.seed(10101)

library(dplyr)
library(brms)
library(ggplot2)
library(tidyr)
library(glue) 
library(caret)
library(tidyverse)
library(randomForest)
library(caTools)
library(bayesplot)
library(glmnet)
library(here)

options(mc.cores = parallel::detectCores()) # Paralellize if possible
options(brms.file_refit = "on_change") # Save the files if the model has changed
```

# Code for the Dyslexia Paper

## Data

Two sources of data are available: The original data, collected through the use of computers, and a second source, collected on tablets with a modified game. (The game had less questions.) No attempt has been made yet to use both datasets; I abandoned my efforts after preliminary testing showed no real difference in performance when I included the secondary dataset. However, I think with a bit more work, perhaps a better imputation method, I could find a way to exploit the information contained in the secondary dataset.

```{r}
data <- read.csv(here("data", "Dyt-desktop.csv"), sep=";", dec=".", header=TRUE)

# Accuracy and Missrate are not accurate in many cases; I fix it here
for (i in 1:32) {
  accuracy_col <- paste0("Accuracy", i)
  missrate_col <- paste0("Missrate", i)
  hits_col <- paste0("Hits", i)
  clicks_col <- paste0("Clicks", i)
  misses_col <- paste0("Misses", i)
  data[[accuracy_col]] <- ifelse(data[[clicks_col]] == 0, 0, data[[hits_col]] / data[[clicks_col]])
  data[[missrate_col]] <- ifelse(data[[clicks_col]] == 0, 00, data[[misses_col]] / data[[clicks_col]])
}

# Changing Gender column to Female column
data$Female <- ifelse(data$Gender == "Male", "Yes", "No")
data$Gender <- NULL

# Factorizing the categorical variables
data$Female <- factor(data$Female, levels = c("No", "Yes"))
data$Nativelang <- factor(data$Nativelang, levels = c("No", "Yes"))
data$Otherlang <- factor(data$Otherlang, levels = c("No", "Yes"))
data$Dyslexia <- factor(data$Dyslexia, levels = c("No", "Yes"))

# Train-test split
sample = sample.split(data[,1], SplitRatio = 0.8)
traintemp = subset(data, sample == TRUE)
testing  = subset(data, sample == FALSE)

# Train-validation split
sample = sample.split(traintemp[,1], SplitRatio = 0.75)
training = subset(traintemp, sample == TRUE)
validation  = subset(traintemp, sample == FALSE)
```

## Bayesian Modelling

```{r}
# BAYESIAN MODEL WITH REG HORSESHOE PRIOR AND UNINFORMATIVE PRIORS

# Specifying formula
formula <- bf(Dyslexia ~ a + b, nl = TRUE) +
  lf(a ~ Clicks1 + Hits1 + Misses1 + Score1 + Accuracy1 + Missrate1 + Clicks2 + Hits2 + Misses2 + Score2 + Accuracy2 + Missrate2 + Clicks3 + Hits3 + Misses3 + Score3 + Accuracy3 + Missrate3 + Clicks4 + Hits4 + Misses4 + Score4 + Accuracy4 + Missrate4 + Clicks5 + Hits5 + Misses5 + Score5 + Accuracy5 + Missrate5 + Clicks6 + Hits6 + Misses6 + Score6 + Accuracy6 + Missrate6 + Clicks7 + Hits7 + Misses7 + Score7 + Accuracy7 + Missrate7 + Clicks8 + Hits8 + Misses8 + Score8 + Accuracy8 + Missrate8 + Clicks9 + Hits9 + Misses9 + Score9 + Accuracy9 + Missrate9 + Clicks10 + Hits10 + Misses10 + Score10 + Accuracy10 + Missrate10 + Clicks11 + Hits11 + Misses11 + Score11 + Accuracy11 + Missrate11 + Clicks12 + Hits12 + Misses12 + Score12 + Accuracy12 + Missrate12 + Clicks13 + Hits13 + Misses13 + Score13 + Accuracy13 + Missrate13 + Clicks14 + Hits14 + Misses14 + Score14 + Accuracy14 + Missrate14 + Clicks15 + Hits15 + Misses15 + Score15 + Accuracy15 + Missrate15 + Clicks16 + Hits16 + Misses16 + Score16 + Accuracy16 + Missrate16 + Clicks17 + Hits17 + Misses17 + Score17 + Accuracy17 + Missrate17 + Clicks18 + Hits18 + Misses18 + Score18 + Accuracy18 + Missrate18 + Clicks19 + Hits19 + Misses19 + Score19 + Accuracy19 + Missrate19 + Clicks20 + Hits20 + Misses20 + Score20 + Accuracy20 + Missrate20 + Clicks21 + Hits21 + Misses21 + Score21 + Accuracy21 + Missrate21 + Clicks22 + Hits22 + Misses22 + Score22 + Accuracy22 + Missrate22 + Clicks23 + Hits23 + Misses23 + Score23 + Accuracy23 + Missrate23 + Clicks24 + Hits24 + Misses24 + Score24 + Accuracy24 + Missrate24 + Clicks25 + Hits25 + Misses25 + Score25 + Accuracy25 + Missrate25 + Clicks26 + Hits26 + Misses26 + Score26 + Accuracy26 + Missrate26 + Clicks27 + Hits27 + Misses27 + Score27 + Accuracy27 + Missrate27 + Clicks28 + Hits28 + Misses28 + Score28 + Accuracy28 + Missrate28 + Clicks29 + Hits29 + Misses29 + Score29 + Accuracy29 + Missrate29 + Clicks30 + Hits30 + Misses30 + Score30 + Accuracy30 + Missrate30 + Clicks31 + Hits31 + Misses31 + Score31 + Accuracy31 + Missrate31 + Clicks32 + Hits32 + Misses32 + Score32 + Accuracy32, center = TRUE) +
  lf(b ~ 0 + Nativelang + Otherlang + Age + Female, cmc = FALSE)

# Specifying priors
priors <- prior(horseshoe(1), class = "b", nlpar = "a")

# Specifying the model
start_time <- Sys.time()

bayesian_model <- brm(formula = formula,
           prior = priors,
           family = bernoulli(link = 'logit'),
           data = training
           )

end_time <- Sys.time()
runtime <- end_time - start_time

writeLines(runtime, "runtime.txt")
saveRDS(bayesian_model, file = "bayesian_model.rds")
```

## Logistic Regression Modelling  

```{r}
# Prepare data for glmnet (training set)
x_train <- model.matrix(Dyslexia ~ . - 1, data = training)
y_train <- training$Dyslexia

# Fit logistic regression with L1/L2 regularization
logistic_model <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 0.5)

# Save the model
saveRDS(logistic_model, file = "logistic_model.rds")
```

## Random Forest Modelling

```{r}
# Build the Random Forest model with weighted classes
rf_model <- train(
  Dyslexia ~ ., 
  data = training, 
  method = "rf",
  ntree = 200, 
  tuneGrid = data.frame(mtry = floor(sqrt(ncol(training) - 1))), 
  weights = ifelse(training$Dyslexia == 1, 0.892, 0.108) # adjust based on dyslexia prevalence, the authors used these values
)

# Save the model
saveRDS(rf_model, file = "rf_model.rds")
```

## Decision Curve Analysis

### Getting models' predictions

```{r}

# We need this for the results
Dyslexia_binary <- ifelse(validation$Dyslexia == "Yes", 1, 0)

# ALL-TREATMENT CASE
all_treatment <- data.frame(
  probabilities = 1,
  outcomes = Dyslexia_binary)

# BAYESIAN MODEL
#bayesian_model <- readRDS("bayesian_model.rds")
bayesian_model_predicted_probabilities <- fitted(bayesian_model, newdata = validation)
bayesian_model_probabilities <- bayesian_model_predicted_probabilities[, 1]

bayesian_model_predictions <- data.frame(
  probabilities = model_1_probabilities,
  outcomes = Dyslexia_binary
)

# LOGISTIC MODEL
#logistic_model <- readRDS("logistic_model.rds")
probabilities <- predict(logistic_model, newdata = validation, type = "response")
probabilities <- unname(probabilities)

log_model_predictions <- data.frame(
  probabilities = probabilities,
  outcomes = Dyslexia_binary
)

# RANDOM FOREST MODEL
#random_forest_model <- readRDS("random_forest_model.rds")
probabilities <- predict(random_forest_model, newdata = validation, type = "prob")[, 2]

rf_model_predictions <- data.frame(
  probabilities = probabilities,
  outcomes = Dyslexia_binary
)

```

### Doing DCA

Dca is done as follows: First I get the models to assign a likelihood for dyslexia for each entry in the dataset being used for predictions. Then I pair the resulting predictions with actual 1-0 outcomes. After this, I make a list data_list out of all these dataframes, collecting all models' predictions in a single entity. I then run the code, which applies my function calculate_net_benefits. This results in plot_data, which contains the net benefits of each model for each threshold. This is then plotted.
 
```{r}
# Getting net benefit values

# List of dataframes to compare, each with probabilities and outcomes
data_list <- list(all_treatment = all_treatment, bayesian_model = bayesian_model_predictions, log_model = log_model_predictions, rf_model = rf_model_predictions) # Add more as needed
thresholds <- seq(0, 1, by = 0.01)

# Calculate net benefits for each dataframe
plot_data <- bind_rows(
  lapply(names(data_list), function(name) {
    df <- data_list[[name]]
    net_benefit_df <- calculate_net_benefits(df = df, probabilities_column = "probabilities", outcomes_column = "outcomes") # Default thresholds
    net_benefit_df$Model <- name # Add a model name for plotting
    net_benefit_df
  })
)

# Filter out NA values resulting from extreme thresholds
plot_data <- plot_data %>% filter(!is.na(NetBenefit))
```

### Plotting

```{r}
# Comparing the models
DCA1 <- ggplot(plot_data, aes(x = Threshold, y = NetBenefit, color = Model)) +
  geom_line(size = 1) +
  labs(
    title = "Net Benefit Curves for\nModelling Paradigm Comparison",
    x = "Threshold",
    y = "Net Benefit"
    ) +
  labs(x = "Decision Threshold", y = "Net Benefit") +
  ylim(0, 0.15) + 
  xlim(0, 0.3) +
   scale_color_manual(
   name = "Modelling\nmethod:",
   labels = c("AllTreatment", "SLR", "RF", "BLR"),
   values = c("skyblue", "forestgreen", "darkslateblue", "firebrick") 
    ) +
  theme(axis.text = element_text(size = 12),
          plot.title = element_text(size = 16, hjust = 0.5),
          panel.grid.major = element_line(color = "grey75", size = 0.5),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          legend.direction = "vertical",
          legend.position = "right",
          legend.background = element_blank(),
          axis.title.y = element_text(size = 14, face = "italic"),
          axis.title.x = element_text(size = 14, face = "italic"),
          panel.border = element_rect(color = "grey50", fill = NA, size = 0.5)
          ) +
  expand_limits(x = 0, y = 0) +
  coord_cartesian(expand = FALSE, clip = "off") +
  guides(color = guide_legend(override.aes = list(size = 3)))

DCA1
```

## What's missing from this notebook

The models are only evaluated with DCA so far. Also, I didn't include any additional diagnostics.